### 2022-06-13 paper.tex
sc-RNA-Seq -> scRNA-seq

Sequencing-based assays with single-cell resolution have offered new means to understand the differences between the cells making up a sample. Single-cell RNA sequencing (sc-RNA-seq) techniques have matured at great pace in recent years, with well developed analysis methodology, and methods to study epigenetics at single-cell resolution are rapidly catching up. 
\blfootnote{\hspace{-.5cm}\raggedright\textsuperscript{*} e-mail:  \texttt{l.kremer@dkfz-heidelberg.de} and\\ \hspace{1cm}\texttt{simon.anders@bioquant.uni-heidelberg.de}}

Here, we discuss bisulfite sequencing, a method to study DNA methylation. In brief, DNA is treated with bisulfite, which converts unmethylated cytosines to uracils which are read as thymine in subsequent PCR, while methylated cytosines are protected from conversion. After sequencing, these conversions allow to determine the methylation status of all cytosines covered by reads \citep{Frommer_1992}. Bisulfite sequencing can also be performed with single-cell resolution \citep{Smallwood_2014} and even in parallel with sc-RNA-Seq \citep{Clark2018}.

@@

This procedure is commonly adapted when working with single-cell DNA-methylation data, so that established methods used in scRNA-Seq can be used.
The task is therefore how to construct a suitable matrix as input for the PCA. A simple and common approach, used for instance by \citet{luo2017single}, is to tile the genome into windows of, say, 100~kb size, and calculate for each cell the average methylation of the DNA within each window. To this end, we identify all CpG sites in the window for which we have coverage with at least one read and can hence call the CpG to be either unmethylated or methylated in the cell. Then we denote as average DNA methylation of this window in a given cell the proportion of observed CpG sites in the window that were found to be methylated (Figure \ref{fig:smoothres}A). This yields a matrix, with one row for each cell and one column for each window from the genome tiling, comprising numbers (methylation fractions) between 0 and 1. This matrix is now subjected to PCA. After the PCA, one can proceed using approaches known from sc-RNA-Seq.

>>>>

Sequencing-based assays with single-cell resolution have offered new means to understand the differences between the cells making up a sample. Single-cell RNA sequencing (scRNA-seq) techniques have matured at great pace in recent years, with well developed analysis methodology, and methods to study epigenetics at single-cell resolution are rapidly catching up. 
\blfootnote{\hspace{-.5cm}\raggedright\textsuperscript{*} e-mail:  \texttt{l.kremer@dkfz-heidelberg.de} and\\ \hspace{1cm}\texttt{simon.anders@bioquant.uni-heidelberg.de}}

Here, we discuss bisulfite sequencing, a method to study DNA methylation. In brief, DNA is treated with bisulfite, which converts unmethylated cytosines to uracils which are read as thymine in subsequent PCR, while methylated cytosines are protected from conversion. After sequencing, these conversions allow to determine the methylation status of all cytosines covered by reads \citep{Frommer_1992}. Bisulfite sequencing can also be performed with single-cell resolution \citep{Smallwood_2014} and even in parallel with scRNA-seq \citep{Clark2018}.

@@

This procedure is commonly adapted when working with single-cell DNA-methylation data, so that established methods used in scRNA-Seq can be used.
The task is therefore how to construct a suitable matrix as input for the PCA. A simple and common approach, used for instance by \citet{luo2017single}, is to tile the genome into windows of, say, 100~kb size, and calculate for each cell the average methylation of the DNA within each window. To this end, we identify all CpG sites in the window for which we have coverage with at least one read and can hence call the CpG to be either unmethylated or methylated in the cell. Then we denote as average DNA methylation of this window in a given cell the proportion of observed CpG sites in the window that were found to be methylated (Figure \ref{fig:smoothres}A). This yields a matrix, with one row for each cell and one column for each window from the genome tiling, comprising numbers (methylation fractions) between 0 and 1. This matrix is now subjected to PCA. After the PCA, one can proceed using approaches known from scRNA-seq.
